---
title: "Basic Tutorial"
author: "Sean Jewell"
date: '2019-10-02'
output: html_document
layout: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```


Basic tutorial
----

In this tutorial, we demonstrate basic use of the ```ChangepointInference``` package. 
Installation instructions are provided [here](software.html). 

First load the package: 
```{r}
require(ChangepointInference)
```

To illustrate the software, we generate a synthetic dataset according to 

$$
\begin{align}
Y_{t} = \mu_{t} + \epsilon_{t}, \quad \epsilon_{t} \overset{\text{iid}}{\sim} \mathrm{N}(0, \sigma^{2}), \quad t=1,\ldots,T,
\end{align}
$$
and assume that $\mu_1,\ldots,\mu_T$ is  piecewise constant, in the sense that  $\mu_{\tau_j+1}=\mu_{\tau_j + 2 } = \ldots = \mu_{\tau_{j+1}}$, $\mu_{\tau_{j+1}} \neq \mu_{\tau_{j+1}+1}$, for $j=0,\ldots,K-1$, where $0 = \tau_{0} < \tau_{1} < \ldots < \tau_{K} < \tau_{K+1} = T$, and where $\tau_1,\ldots,\tau_K$ represent the true changepoints.  


```{r}
set.seed(1)
mu <- rep(c(1, 3, -2), each = 100)
dat <- mu + rnorm(length(mu))
plot(dat, cex = 1, ylab = "y", xlab = "t")
```

#### Changepoint estimation

* Estimate changepoints using $\ell_0$ segmentation and with tuning parameter $\lambda = 4$:
```{r}
lam <- 4
fit <- changepoint_estimates(dat, "L0", lam)
print(fit)
```

The ```fit``` object contains model fit information:
```{r}
str(fit)
```

Importantly, the estimated changepoints are 
```{r}
fit$change_pts
```

* Estimate changepoints using binary segmentation with $K = 2$ changepoints:

```{r}
K <- 2
fit <- changepoint_estimates(dat, "BS", K)
print(fit)
```

The ```fit``` object contains model fit information:
```{r}
str(fit)
```

Importantly the sorted changepoints, the order changepoints were estimated, and the sign of the change in mean due to a changepoint (in same order as ```fit$ordered_change_pts```) are:
```{r}
fit$change_pts
fit$ordered_change_pts
fit$change_pt_signs
```

In particular, we note that in this example, the first changepoint is estimated at time point $200$ and at this point the sign of the change in mean is negative. This is expected as the mean of $y_{201:300}$ is less than the mean of $y_{1:200}$.  


#### Changepoint Inference

In this section we demonstrate how to use our software to obtain $p$-values for the following test statistics and conditioning sets. For the fixed window tests, we take the window size $h = 10$. 

* Type = 'L0-fixed': 
    $$
    \text{Pr}(|\phi| \geq |\nu^\top y| | \hat{\tau}_j \in \mathcal{M}(y'(\phi))),
    $$
    for fixed $\nu$.
    
    ```{r}
    h <- 10
    fit_inference <- changepoint_inference(dat, 'L0-fixed', lam, window_size = h, sig = 1)
    knitr::kable(data.frame(estimated_changepoints = fit_inference$change_pts, pvals = fit_inference$pvals))
    ```


* Type = 'BS-fixed': 
    $$
    \text{Pr}(|\phi| \geq |\nu^\top y| | \hat{\tau}_j \in \mathcal{M}(y'(\phi))),
    $$
    for fixed $\nu$.
    
    ```{r}
    fit_inference <- changepoint_inference(dat, 'BS-fixed', K, window_size = h, sig = 1)
    knitr::kable(data.frame(estimated_changepoints = fit_inference$change_pts, pvals = fit_inference$pvals))
    ```

* Type = 'BS-adaptive-M-O-D':
    $$
    \text{Pr}(|\phi| \geq |\nu^\top y| | \mathcal{M}(y) = \mathcal{M}(y'(\phi)), \mathcal{O}(y) = \mathcal{O}(y'(\phi)), 
    \Delta(y) =
     \Delta(y'
    (\phi)),
    $$
    for adaptive $\nu$.
    
    ```{r}
    fit_inference <- changepoint_inference(dat, 'BS-adaptive-M-O-D', K, sig = 1)
    knitr::kable(data.frame(estimated_changepoints = fit_inference$change_pts, pvals = fit_inference$pvals))
    ```

* Type = 'BS-adaptive-M-O':
    $$
    \text{Pr}(|\phi| \geq |\nu^\top y| | \mathcal{M}(y) = \mathcal{M}(y'(\phi)), \mathcal{O}(y) = \mathcal{O}(y'(\phi))),
    $$
    for adaptive $\nu$. 
    
    ```{r}
    fit_inference <- changepoint_inference(dat, 'BS-adaptive-M-O', K, sig = 1)
    knitr::kable(data.frame(estimated_changepoints = fit_inference$change_pts, pvals = fit_inference$pvals))
    ```

* Type = 'BS-adaptive-M':
    $$
    \text{Pr}(|\phi| \geq |\nu^\top y| | \mathcal{M}(y) = \mathcal{M}(y'(\phi))), 
    $$
    for adaptive $\nu$.
    
    ```{r}
    fit_inference <- changepoint_inference(dat, 'BS-adaptive-M', K, sig = 1)
    knitr::kable(data.frame(estimated_changepoints = fit_inference$change_pts, pvals = fit_inference$pvals))
    ```